--- 
layout: post
title: Ranking in R
published: true
tags: R
type: post
status: process
---

## How to implement Yelp-style ranking in R

Would you rather go to a restaurant that has 5 stars but only one review or a place with 4.9 stars on 1,000 reviews? Even though the first restaurant technically has a higher rating, the second restaurant is probably the better choice. [Evan Miller](http://www.evanmiller.org/how-not-to-sort-by-average-rating.html) posted a how-to guide back in 2009 this kind of ranking. Miller has code examples in Ruby, SQL, and even Excel, but not R. Here's how to implement it in the best statistical programming language.

First we generate some dummy data. For simplicity's sake, let's treat a rating of 0 as a bad review and a rating of 1 as a good review.

```{r, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(tidyverse))

x <-
  tibble::tibble(
  product = sample(LETTERS, size = 1000, replace = TRUE),
  rating = sample(c(0, 1), size = 1000, replace = TRUE)
  )
```

### Warning: math ahead

Now for the actual ranking. There's a lot of math involved, but the equation is fortunately quite easy to implement with some help from the [tidyverse](https://www.tidyverse.org). Check out Evan's original article for details on the algorithm.

```{r, message=FALSE,echo=TRUE}

accurate_rankings <- x %>%
  group_by(product) %>%
  summarise(ci_lower_bound = ((sum(rating == 1) + 1.9208) / (sum(rating == 1) + sum(rating == 0)) -
  1.96 * sqrt((sum(rating == 1) * sum(rating == 0)) / (sum(rating == 1) + sum(rating == 0)) + 0.9604) /
  (sum(rating == 1) + sum(rating == 0))
  ) / (1 + 3.8416 / (sum(rating == 1) + sum(rating == 0)))) %>%
  arrange(desc(ci_lower_bound))
```

